\LARGE \bf The Battle for Filter Supremacy: A Comparative Study of the Multi-State Constraint Kalman Filter and the Sliding Window Filter

### 2014-12-31 Report/project_report.tex
vspace to the rescue!
Renumbered figs
Fixed numbering
Final touches

\begin{equation}
\pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} = \frac{1}{z^2} \bbm f_u z & 0 & - f_u x \\  0 & f_v z & -f_v y \ebm.
\end{equation}
The weight matrix $\Vec T$ is given by:
\begin{equation}
\Vec{T} := \text{diag} \left\{ \Vec{T}_1,\hdots,\Vec{T}_K \right\}
\end{equation}
where
\begin{equation}
\Vec{T}_k := \bbm \Vec H_{\Vec w, k} \Vec Q_k \Vec H_{\Vec w, k}^T & \Vec 0 \\
\Vec 0 & \Vec G_{\Vec n, k} \Vec R_k \Vec G_{\Vec n, k}^T \ebm,
\end{equation}
and $\Vec R_k$, $\Vec Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).

@@

\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/NST_trans}
        \label{fig:nullspace:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/NST_rot}
        \label{fig:nullspace:rot_err}
    }

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth}
    \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
    \label{fig:trajectory_groundtruth}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/nVisible_500_1000}
    \caption{Number of visible features in the interval $k\in[500,1000]$. Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
    \label{fig:nVisible_int1}
\end{figure}

@@

\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
        \label{fig:comp_1:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
        \label{fig:comp_1:rot_err}
    }

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/nVisible_1215_1715}
    \caption{Number of visible features in the interval $k\in[1215,1715]$. Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
    \label{fig:nVisible_int2}
\end{figure}

\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
        \label{fig:comp_2:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
        \label{fig:comp_2:rot_err}
    }

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
    \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
    \label{fig:trajectory_groundtruth_100lessnoisy}
\end{figure}

@@

Note that the MSCKF's performance improves as the number of features increases, while the SWF's performance is not significantly affected by increasing feature count.
This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
This may be due to the fact that the MSCKF updates its state whenever a feature goes out of view and does not associate tracks corresponding to the same feature.
If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF benefits from more high-quality motion constraints.
Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.

\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
        \label{fig:comp_3:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
        \label{fig:comp_3:rot_err}
    }

@@

Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the synthetic datasets.
As expected from Figure \ref{fig:comp_3}, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.

@@

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $k\in[1215,1715]$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Position ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rotation ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
                    & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF (25 Poses)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
                    & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions} \label{sec:conclusions}
Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
We stress that our inertial data was obtained from a high-quality IMU and has been sanitized to account for gravity, biases, and integration error.
Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent.

>>>>

\begin{equation}
\pd{\Vec g}{ \Vec p}\Bigr|_{\bar{\Vec p}_{C_k}^{f_j C_k}} = \frac{1}{z^2} \bbm f_u z & 0 & - f_u x \\  0 & f_v z & -f_v y \ebm.
\end{equation}
The weight matrix $\Vec T$ is given by
\begin{equation}
\Vec{T} := \text{diag} \left\{ \Vec{T}_1,\hdots,\Vec{T}_K \right\}
\end{equation}
where
\begin{equation}
\Vec{T}_k := \bbm \Vec H_{\Vec w, k} \Vec Q_k \Vec H_{\Vec w, k}^T & \Vec 0 \\
\Vec 0 & \Vec G_{\Vec n, k} \Vec R_k \Vec G_{\Vec n, k}^T \ebm,
\end{equation}
and $\Vec R_k$, $\Vec Q_k$ are observation and motion model noise covariances, and $\Vec G_{\Vec n, k}$, $\Vec H_{\Vec w, k}$ are the observation and motion model Jacobians with respect to the noise.
The update is performed repeatedly until convergence (in this work we use the criterion $|\delta \Vec{x}^*| < 10^{-3} $).

@@

\setcounter{figure}{2}
\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/NST_trans}
        \label{fig:nullspace:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/NST_rot}
        \label{fig:nullspace:rot_err}
    }

@@

\setcounter{figure}{1}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth}
    \caption{Vicon ground truth for sensor head motion and feature positions in the ``Starry Night'' dataset.}
    \label{fig:trajectory_groundtruth}
\end{figure}

@@

\setcounter{figure}{3}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/nVisible_500_1000}
    \caption{Number of visible features in the interval $k\in[500,1000]$. Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
    \label{fig:nVisible_int1}
\end{figure}

@@

\setcounter{figure}{5}
\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Trans}
        \label{fig:comp_1:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-500-1000-Rot}
        \label{fig:comp_1:rot_err}
    }

@@

\setcounter{figure}{4}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/nVisible_1215_1715}
    \caption{Number of visible features in the interval $k\in[1215,1715]$. Red points indicate that fewer than three features were visible, while green points indicate at least three features were visible.}
    \label{fig:nVisible_int2}
\end{figure}

\setcounter{figure}{6}
\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Trans}
        \label{fig:comp_2:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/6-Way-Comparison-1215-1715-Rot}
        \label{fig:comp_2:rot_err}
    }

@@

\setcounter{figure}{8}
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figs/trajectory_groundtruth_100lessnoisy}
    \caption{Modified ``Starry Night'' dataset with 100 synthetic features. The sensor trajectory is identical to the original dataset.}
    \label{fig:trajectory_groundtruth_100lessnoisy}
\end{figure}

@@

Note that the MSCKF's performance improves as the number of features increases, while the SWF's performance is not significantly affected by increasing feature count.
This result indicates that the MSCKF is much more sensitive to feature density than the SWF.
This may be due to the fact that the MSCKF updates its state whenever a feature goes out of view and does not associate tracks corresponding to the same feature.
If observations of a given feature are frequently interrupted, each track will not constrain the full set of poses from which the feature is visible.
As feature density increases, the number of long feature tracks is likely to increase, and so the MSCKF benefits from more high-quality motion constraints.
Since the SWF always associates all observations of each feature in a given window, it is not sensitive to the contiguity of the feature observations.
\setcounter{figure}{7}
\begin{figure*}
    \centering
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Trans}
        \label{fig:comp_3:trans_err}
    }
    ~
    \subfigure {
        \includegraphics[width=0.45\textwidth]{figs/RMSE-Comparison-Feat-Density-1215-1715-Rot}
        \label{fig:comp_3:rot_err}
    }

@@

Table \ref{tab:rmse_nees} provides summary statistics of each algorithm's performance on the interval shown in Figure \ref{fig:comp_3}.
As expected, the SWF outperforms the MSCKF in terms of average RMSE on both rotation estimation and position estimation.
It is worth noting, however, that, on average, the MSCKF achieves low Normalized Estimation Error Squared (NEES) values, which indicates that the filter scores well on consistency if not on accuracy.
The average NEES values for the SWF are substantially higher because the SWF treats each window independently and has no built-in mechanism for propagating uncertainty from window to window (\cite{Sibley:2010:JFR} describes how this can be remedied by marginalizing old poses, a technique which was not implemented in this work).
The accumulated drift error in the SWF quickly escapes the estimator's reported uncertainty envelope, and so this form of the SWF may be considered inconsistent in the absence of a known map.

@@

\begin{table}
    \centering
    \caption{Comparison of Average Root Mean Squared Error (ARMSE), Average Normalized Estimation Error Squared (ANEES), and Compute Time of IMU integration, MSCKF, and SWF on synthetic datasets for the interval $k\in[1215,1715]$.}
    \begin{threeparttable}
    \begin{tabular}{llccc} 
                    &                   & \multicolumn{3}{c}{Feature Count} \B      \\ \cline{3-5}
                    &                   & 40        & 60        & 100       \T\B    \\ \hline
        IMU Only    & Position ARMSE    & 0.3679    & 0.3679    & 0.3679    \T      \\
                    & Rotation ARMSE    & 0.1452    & 0.1452    & 0.1452            \\
                    & ANEES             & 0.2850    & 0.2850    & 0.2850            \\
                    & Compute Time\tnote{\dag}      & 8.90 s    & 8.90 s    & 8.90 s    \B      \\ \hline
        MSCKF (20-100)       & Position ARMSE    & 0.2672    & 0.2550    & 0.2304    \T      \\
                    & Rotation ARMSE    & 0.1378    & 0.1247    & 0.0952            \\
                    & ANEES             & 10.18     & 12.03     & 16.76             \\
                    & Compute Time\tnote{\dag}     & 12.19 s   & 14.64 s   & 20.58 s   \B      \\ \hline
        SWF (25)         & Position ARMSE    & 0.1750    & 0.1687    & 0.1755    \T      \\
                    & Rotation ARMSE    & 0.0495    & 0.0377    & 0.0481            \\
                    & ANEES             & 2280      & 2093      & 2013              \\
                    & Compute Time\tnote{\dag}     & 114.3 s   & 175.9 s   & 245.3 s   \B      \\ \hline
    \end{tabular}
    \begin{tablenotes}
        \item [\dag] Running MATLAB 2014b on a MacBook Pro Retina (11,3) with a 2.3GHz Intel Core i7 processor and 16 GB of DDR3L RAM.
    \end{tablenotes}
    \end{threeparttable}
    \label{tab:rmse_nees}
    \vspace{-0.4cm}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.1cm}
\section{Conclusions} \label{sec:conclusions}
Based on our experimental results, we conclude that neither the SWF nor the MSCKF is more effective than pure IMU integration in environments with few features.
We stress that our inertial data was obtained from a high-quality IMU and has been sanitized to account for gravity, biases, and integration error.
Using raw data from a consumer-grade IMU, we expect that pure IMU integration would have performed much worse, and that the benefit of the SWF and MSCKF would have been more apparent.
